---
params:
  files: ""
title: "Benchmark single cell long-read"
subtitle: "single cell metrics visualization"
author: "Ali Hamraoui & Audrey Onfroy"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    code_folding: hide
    code_download: true
    toc: true
    toc_float: true
    number_sections: false
---

<style>
body {
text-align: justify}

.flex-container {
  display: flex;
  justify-content: start;
}

.flex-container > div {
  margin-right: 10px; /* Espacement entre les boutons */
}

button {
  background-color: #1f77b4; /* Couleur de fond bleue Plotly */
  border: none;
  color: white;
  padding: 10px 24px;
  text-align: center;
  text-decoration: none;
  display: inline-block;
  font-size: 16px;
  margin: 4px 2px;
  cursor: pointer;
  border-radius: 8px; /* Bordures arrondies */
  box-shadow: 0 4px 8px 0 rgba(0,0,0,0.2), 0 6px 20px 0 rgba(0,0,0,0.19); /* Ombre pour un effet en relief */
}

button:hover {
  background-color: #ff7f0e; /* Couleur de fond orange au survol, autre couleur Plotly */
}
</style>

<script type="text/javascript">
function toggleImage(id) {
  var x = document.getElementById(id);
  if (x.style.display === "none") {
    x.style.display = "block";
  } else {
    x.style.display = "none";
  }
}
</script>

<!-- Automatically computes and prints in the output the running time for any code chunk -->
```{r, echo=FALSE}
# https://github.com/rstudio/rmarkdown/issues/1453
hooks = knitr::knit_hooks$get()
hook_foldable = function(type) {
  force(type)
  function(x, options) {
    res = hooks[[type]](x, options)
    
    if (isFALSE(options[[paste0("fold_", type)]])) return(res)
    
    paste0(
      "<details><summary>", "show", "</summary>\n\n",
      res,
      "\n\n</details>"
    )
  }
}
knitr::knit_hooks$set(
  output = hook_foldable("output"),
  plot = hook_foldable("plot")
)
```

<!-- Set default parameters for all chunks -->
```{r, setup, include = FALSE}
set.seed(1)
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fold_output = FALSE,
                      fold_plot = FALSE,
                      
                      fig.align = 'center',
                      fig.width = 3,
                      fig.height = 3)
```

The goal of this script is to generate a data object for sample `r params$sample_name`.

* removal of cells based on quality control metrics

```{r locations, message=FALSE, warning=FALSE, paged.print=FALSE}
load("cell_markers.rda")
load("color_markers.rda")
source("common_imports.R")
source("clustering.R")
```

## Data

```{r}
files_list <- params$files
files_list <- strsplit(files_list, " ")[[1]]
files_list
```


#```{r load_Sobj}
#obj.lsit <- c("nanopore.Sobj", "integrated.Sobj", "illumina.Sobj")
#for (obj in obj.lsit){
#    idx <- grep(obj, files_list)
#    for (file in files_list[idx]){
#        data_name <- strsplit(file, split ="_")[[1]][1]
#        Sobj.tmp <- readRDS(file)
#        assign(paste0(obj, ".", data_name), Sobj.tmp)
#        }
#}
#rm(Sobj.tmp)
#```

```{r load_data}
for (dt in c("cLISI", "iLISI", "ARI")){
    li.tmp <- list()
    idx <- grep(dt, files_list)
          for (file in files_list[idx]){
              data_name <- strsplit(file, split ="_")[[1]][1]
              li.tmp[[data_name]] <- as.data.frame(readRDS(file))
              }
    assign(dt, li.tmp)
}
```

```{r combine_LISI}
df.cols <- c("data","LISI","workflow")

iLISI <- ldply(iLISI, rbind)
colnames(iLISI) <-  df.cols

cLISI <- ldply(cLISI, rbind)
colnames(cLISI) <-  df.cols

df.lisi <- list("cLISI" = cLISI, "iLISI" = iLISI)

df.lisi <-  ldply(df.lisi, rbind)
rm(iLISI,cLISI)

df.cols <- c("data","ARI","workflow")

ARI <- ldply(ARI, rbind)
colnames(ARI) <-  df.cols

```

```{r fig.height=2, fig.width=3}
boxplot_lisi <- function(lisi, title){
df.lisi[df.lisi$.id==lisi ,] %>%
  ggplot(aes(fill=data, y=LISI-1, x=workflow))+
    geom_boxplot(width = 0.5, outlier.shape = NA) +
    theme_classic()+
    labs(x = "", y = lisi, title=title) +
    theme(axis.text.x =  element_text(colour = "black"),
          axis.text = element_text(face="bold"),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank()) 
}
```

```{r fig.height=4, fig.width=8}
bar_metric <- function(df, title, desc=F){
    df %>%
    ggplot(aes(x = workflow, y = ARI, fill = data)) +
              geom_bar(stat = "identity", position=position_dodge()) +
              theme_classic()+
              labs(x = "", y = "", title=title) +
              theme(axis.text.x =  element_text(colour = "black"),
                   axis.text = element_text(face="bold"),
                   axis.ticks.x = element_blank(),
                   axis.title.x = element_blank())
}
```

```{r fig.height=4.5, fig.width=14}
p1 <- ggarrange(boxplot_lisi("cLISI", "Cell type LISI")+coord_flip(),
          boxplot_lisi("iLISI", "Integration LISI")+coord_flip(), 
          bar_metric(ARI, "ARI") + coord_flip(),
          ncol = 3, common.legend =T, legend="bottom")
p1 
```

```{r fig.height=4, fig.width=6}
pdf(file = 'quanti_metrics.pdf',
  width =13,
  height = 4.5) 
p1 
dev.off()
```

```{r create_summary_df}
df_summary <- data.frame()
for (tool in unique(ARI$workflow)){
    
    mean_ARI <- median(ARI[ARI$workflow == tool,]$ARI)
    median_cLisi <- median(df.lisi[df.lisi$workflow == tool & df.lisi$.id == 'cLISI' ,]$LISI)
    median_iLisi <- median(df.lisi[df.lisi$workflow == tool & df.lisi$.id == 'iLISI' ,]$LISI)
    
    df <- data.frame(tool=tool, mean_ARI=mean_ARI, median_cLisi=median_cLisi, median_iLisi=median_iLisi)
    df_summary <- rbind(df, df_summary)
}
```

```{r save_summary_df}
write_csv(df_summary, "cellTypeMetrics.csv")
```

```{r session}
sessionInfo()
```
