---
params:
  files: ""
title: "NF1 long-read"
subtitle: "Visualize QC"
author: "Ali Hamraoui & Audrey Onfroy"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    code_folding: show
    code_download: true
    toc: true
    toc_float: true
    number_sections: false
---

<style>
body {
text-align: justify}
</style>

<!-- Automatically computes and prints in the output the running time for any code chunk -->
```{r, echo=FALSE}
# https://github.com/rstudio/rmarkdown/issues/1453
hooks = knitr::knit_hooks$get()
hook_foldable = function(type) {
  force(type)
  function(x, options) {
    res = hooks[[type]](x, options)
    
    if (isFALSE(options[[paste0("fold_", type)]])) return(res)
    
    paste0(
      "<details><summary>", "show", "</summary>\n\n",
      res,
      "\n\n</details>"
    )
  }
}
knitr::knit_hooks$set(
  output = hook_foldable("output"),
  plot = hook_foldable("plot")
)
```

<!-- Set default parameters for all chunks -->
```{r, setup, include = FALSE}
set.seed(1)
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fold_output = FALSE,
                      fold_plot = FALSE,
                      
                      fig.align = 'center',
                      fig.width = 15,
                      fig.height = 8)
```

The goal of this script is to generate a data object for sample `r params$sample_name`.

* removal of cells based on quality control metrics

```{r libs}
library(ggpubr)
source("common_imports.R")
source("QC.R")
```

## Data
We used the gene expression matrix to count the number of expressed genes per cell. Expressed gene -> one or more IMUs detected from this gene.

```{r list_files}
files_list <- params$files
files_list <- strsplit(files_list, " ")[[1]]
files_list
```


```{r data}
for (dt in c("QC.per.cell", "total_UMI")){
    li.tmp <- list()
    idx <- grep(dt, files_list)
          for (file in files_list[idx]){
              data_name <- strsplit(file, split ="_")[[1]][1]
              print(data_name)
              li.tmp[[data_name]] <- as.data.frame(readRDS(file))
              }
    assign(dt, li.tmp)
}

data.QC <- ldply(QC.per.cell, rbind)
```

```{r correlation_box}
box.cor <-  data.QC %>%
  ggplot(aes(fill=data, y=correlation, x=workflow))+
    geom_boxplot(position = position_dodge2(preserve = "single"), width = 0.5, outlier.shape = NA) +
    theme_classic()+
    labs(x = "", y = "Pearson's correlation", title ="Gene exp correlation with SR" ) +
    theme(axis.text.x =  element_text(colour = "black"),
          axis.text = element_text(face="bold"),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank())
```


### Number of detected UMI per cell

**Total Number of UMI per cell for common cells for different preprocessing workflows against Cell Ranger on the PromethION dataset. The Pearson correlation coefficient is ploted**

```{r rbind_QC}
count.per.cell <- do.call(rbind, QC.per.cell)
```


```{r umi_box}
box.umi <- count.per.cell %>%
  ggplot(aes(fill=data, y=umi, x=workflow))+
    ylim(c(0,25000))+
    geom_boxplot(position = position_dodge2(preserve = "single"), width = 0.5, outlier.shape = NA) +
    theme_classic()+
    labs(x = "", y = "UMI count", title ="No. of UMI per cell") +
    theme(axis.text.x =  element_text(colour = "black"),
          axis.text = element_text(face="bold"),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank())
```

```{r genes_box}
box.genes <- count.per.cell %>%
  ggplot(aes(fill=data, y=gene, x=workflow))+
    ylim(c(0,5000))+
    geom_boxplot(position = position_dodge2(preserve = "single"), width = 0.5, outlier.shape = NA) +
    theme_classic()+
    labs(x = "", y = "Gene count", title ="No. of genes per cell") +
    theme(axis.text.x =  element_text(colour = "black"),
          axis.text = element_text(face="bold"),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank())
```

```{r umi_bar}
df_UMItotal <- rbind(total_UMI$MinION, total_UMI$PromethION, total_UMI$Spatial, total_UMI$ONT)
bar.total <-  df_UMItotal %>%
  ggplot(aes(fill=data, y=total_UMI, x=pipe))+
    geom_bar(stat="identity", position = position_dodge2(preserve = "single")) +
    theme_classic()+
    labs(x = "", y = "", title ="No. total of UMI") +
    scale_y_continuous(labels = scales::number_format(scale = 1e-6, suffix = "M")) +
    theme(axis.text.x =  element_text(colour = "black"),
          axis.text = element_text(face="bold"),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank())
```

```{r viz, fig.height=5, fig.width=15}
ggarrange(box.umi+coord_flip(), box.genes+coord_flip(), box.cor+coord_flip(), bar.total+coord_flip(), ncol = 4, common.legend =T, legend="bottom")
```

```{r log_transform }
count.per.cell$log.umi <- log(count.per.cell$umi)
count.per.cell$log.gene <- log(count.per.cell$gene)

count.per.cell$log.umi.illumina <- log(count.per.cell$umi.illumina )
count.per.cell$log.gene.illumina <- log(count.per.cell$gene.illumina)
```

```{r umi_per_cell, fig.height=6, fig.width=14}


scater.upc <- ggscatter(count.per.cell, x = "log.umi.illumina", y = "log.umi", size = 0.2,
                                                    color = "data") +
          facet_wrap(~workflow) + 
          geom_abline(intercept = 0, slope = 1, color = "gray") +
          stat_cor(aes(color = data, label = after_stat(r.label)), label.y = c(8.7,10,11.3)) +
          facet_grid("UMI per cell" ~ workflow, scales="free_y") +
          xlab("") +
          ylab("Long read") + 
          scale_fill_brewer(palette = "Dark2")+
          theme(axis.text.x = element_text(size = 8),
                axis.text.y = element_text(size = 8),
                strip.background = element_rect(fill = "transparent"),
                strip.text.x = element_text(size = 8)
        )

scater.gpc <- ggscatter(count.per.cell, x = "log.gene.illumina", y = "log.gene", size = 0.2,
                                                    color = "data") +
          facet_wrap(~workflow) +
          geom_abline(intercept = 0, slope = 1, color = "gray") +
          stat_cor(aes(color = data, label = after_stat(r.label)), label.y = c(8,9,10)) +

          facet_grid( "GENE per cell" ~ workflow, scales="free_y") +
          xlab("Short read") +
          ylab("Long read") +
          theme(axis.text.x = element_text(size = 8),
                axis.text.y = element_text(size = 8),
                strip.background = element_rect(fill = "transparent"),
                strip.text.x = element_blank()
          )

ggarrange(scater.upc, scater.gpc, nrow=2, common.legend = TRUE)
```


```{r save_pdf}
pdf(file = "QC.pdf", onefile=FALSE,
    width = 15, height = 5)

ggarrange(box.umi+coord_flip(), box.genes+coord_flip(), box.cor+coord_flip(), bar.total+coord_flip(), ncol = 4, common.legend =T, legend="bottom")
dev.off()

pdf(file = "scatters.pdf", onefile=FALSE,
    width = 14, height = 6)

ggarrange(scater.upc, scater.gpc, nrow=2, common.legend = TRUE)

dev.off()
```

```{r create_summary_df}
df_summary <- data.frame()
for (tool in unique(data.QC$workflow)){
    
    cor.umi <- c()
    cor.gene <- c()
      for (dt in unique(count.per.cell$data)){
          df <- count.per.cell[count.per.cell$workflow == tool & count.per.cell$data == dt ,]
          if (nrow(df) > 50) {
                  cor.umi <- c(cor.umi, cor(df$umi, df$log.umi.illumina, ))
                  cor.gene <- c(cor.gene, cor(df$gene, df$log.gene.illumina, ))
                  }
        }
    
    UMI_per_cell_corr <- median(cor.umi)
    Gene_per_cell_corr <- median(cor.gene)
    
    exp_corr <- median(data.QC[data.QC$workflow == tool, ]$correlation)
    
    df <- data.frame(tool=tool, cor.gene=Gene_per_cell_corr, cor.umi=UMI_per_cell_corr, cor.exp=exp_corr)
    df_summary <- rbind(df, df_summary)
}
```

```{r save_summary_df}
write_csv(df_summary, "short_read_correlation.csv")
```

```{r session}
sessionInfo()
```


